### 2018-05-11 General WLD combinatorics.tex
Incorporated notes from May 10 discussion

\begin{dfn}\label{WLdfn}
A Wilson loop diagram is given by the following data: a cyclicly ordered set $V$, and $k$ pairs $\{p_r = (i_r, j_r)\}_{r=1}^k$ ordered such that $i_r +1 < j_r$. \end{dfn}

We depict this data as a convex polygon, with vertices labeled by $V$ (preserving the cyclic ordering), and $k$ wavy lines in the interior of the diagrams. The pair $p_r$ defines a wavy line from the $i_r^{th}$ edge (between $i_r$ and $i_r+1$) and the $j_r^{th}$ edge. These are called propagators.  The condition on $i_r$ and $j_r$ means that the propagator does not go between adjacent edges. Let $\cP = \{p_r\}_{r=1}^k$ be the set of propagators. Then we write \bas W = (\cP, V) \;.\eas

@@

\begin{dfn}\label{equivdfn}
There is an equivalence relationship on the set of admissible Wilson loops diagrams generated by the following binary relation: $W = (\cP, n) \sim W'= (\cP', n)$ if
\begin{enumerate}
\item There exist two different exact subdiagrams, $(P, V(P))$ and $(P', V(P'))$ of $W$ and $W'$ respectively such that $V(P) =  V(P')$.
\item The complementary subdiagrams are identical: $(\cP \setminus P, V(P)^c) = (\cP' \setminus P', V(P')^c)$.
\end{enumerate}
\end{dfn}

@@

The diagrams $W_1 \sim W_2$ because $(\{(5,8), (5,7), \{5,6,7,8,1\})$ and $(\{(5,8), (7,8), \{5,6,7,8,1\})$ are the corresponding differing subdiagrams. Furthermore, there is the equivalence $W_2 \sim W_3$ due to the exact subdiagrams $(\{(1,4), (3,4)\}, \{1,2,3,4,5\})$ and $(\{(1,4), (1,3)\}, \{1,2,3,4,5\})$. This forces an equivalence between $W_1$ and $W_3$, even though one cannot partition the propagators of each into a an exact subdiagram (that may vary between the diagrams) and a complement that is fixed.  \end{eg} \note{Did you change $\sim$ to $\equiv$? Or was I being flaky? If you prefer different notation, lets discuss.}

@@

\begin{dfn}
  Let $W = (\cP, n)$ be an admissible Wilson loop diagram.  The \emph{polygon partition} associated to $W$, denotes $\tau(W)$ is defined as follows.
  \begin{itemize}
  \item The vertices of $\tau(W)$ correspond to the edges of $W$. 
  \item Labeling the edges of $\tau(W)$ with the edge number gives a cyclic order to the vertices. Connecting consecutive vertices gives a graph theoretic cycle called the polygon of $\tau(W)$.
  \item Each propagator of $W$ defines a chord edge of $\tau(W)$ for each propagator of $W$.  If $p= (i,j) \in \cP$, it defines a chord connecting the verices $i$ and $j$
  \end{itemize}
\end{dfn}

\begin{lem}\label{tausimpleplanarlem}
If $W = (\cP, n)$ is an admissible Wilson loop diagram, $\tau(W)$ is a simple planar graph with cyclically ordered vertices. It is embedded such that the vertices all like on the infinite face.
\end{lem}

\begin{proof}
Since the vertices of $\tau(W)$ are labeled by the edges of $W$, which are cyclically ordered, this gives an ordering to the vertices. Since $W$ is admissible, no pairs of propagators cross. Thereofore, it is a planar embedding. There are no propagators $p = (i, i+1)$. Therefore, there is exactly one edge connecting any two adjacent edges of $\tau(W)$. Similarly, there are not two propagators such that both $p$ and $q$ start and end at the edges $i$ and $j$. Therefore, no other two vertices of $\tau(W)$ can be connected by more than one edge. Finally, the embedding of $\tau(W)$ is induced from embedding of the graph $W$. 
 \end{proof}

\hlfix{For an admissible $W$ no propagators cross and so no chord edges of $\tau(W)$ cross.  Thus, in graph theoretic language, we can view $\tau(W)$ as a planar embedding of a
graph with no cut vertices, where all vertices lie on the infinite face, along with a distinguished start vertex and direction around the infinite face.  From this viewpoint the polygon of $tau(W)$ is the boundary of the infinite face.}{I've smushed this together a bit in the prev. lem. Tell me if I've missed anything. Its a lemma for emphasis, not difficulty}

@@

\begin{dfn}
  Let $W$ be an admissible Wilson loop diagram and $\tau(W)$ its polygon partition. A \emph{triangulated piece} of $\tau(W)$ is a 2-connected subgraph of $\tau(W)$ which is a triangulation. We will take the convention that a subgraph consisting of {\color{red}two non-adjascent} vertices joined by an edge  is also a triangulated piece, called a \emph{trivial} triangulated piece.
\end{dfn}

@@

\begin{lem} \label{decompositionlem}
  For $W$ an admissible Wilson loop diagram, the polygon partition $\tau(W)$ has a unique decomposition into maximal triangulated pieces, and boundary edges. 
\end{lem}

@@

\begin{proof}
We begin by giving an algorithm for the decomposition, then prove its uniqueness.

Let $T(W)$ be a the dual graph of $\tau(W)$ with the vertex corresponding the to the infinite face removed. In other words, place a vertex on each finite face of $\tau(W)$. These vertices are connected if there is a chord edge separating the edges. Furthermore, if the face is bounded by an edge of the polygonal cycle of $\tau(W)$, the corresponding vertex gets a leaf edge for each such boundary. Since, by Lemma \ref{tausimpleplanarlem}, $\tau(W)$ is a simple planar graph, this \hlfix{$T(W)$ is an uniquely defined tree.}{Why do I know it is a tree? }

By construction, each face of $\tau(W)$ is a polygon. Therefore, each vertex of $T(W)$ is at least 3 valent (one for each edge of the polygon). If split each vertex of $T(W)$ if it is strictly greater than $3$ valent. (In other words, we fail to split exactly when the face is a triangle). The connected components of $T(W)$ correspond to the decomposition of $\tau(W)$ into maximal triangulated pieces. The the vertices of the forest thus obtained have valence 1, corresponding to root or leaf vertices, or 3, corresponding to triangulated faces. Subtrees with no trivalent vertices correspond to either edges of the boundary of $\tau(W)$, if they were originally leaves of $T(W)$, or to maximal trivial triangulations. \hlfix{Splitting at all the faces that are not triangles ensures maximality of the decomposition.}{more words?}

\hlfix{To see uniqueness, consider the uniqueness of $T(W)$. If there were two different decompositions of $\tau(W)$, this would imply that there are two different ways to split a vertex of $T(W)$, which implies the n0n-uniqueness of the valency of the vertices of $T(W)$, which contradicts the uniqueness of $T(W)$.}{Your proof was much more complicated. What am I missing?}
\end{proof}

@@

  Since $W$ is admissible, by lemma \ref{} $\tau(W)$ is a simple graph. Furthermore, $t$ is not all of $\tau(W)$.
  Let $t$ be a triangulated piece of the decomposition of $\tau(W)$ given in Lemma \ref{decomplem} and suppose $t$ has $n>2$ vertices.  We want to count how many edges of $t$ are not on the outer face of $\tau(W)$.  Consider the intersection of $t$ with the outer face of $\tau(W)$.  This is a possibly disconnected subgraph of the polygon of $\tau(W)$ with $n$ vertices, call it $S$.  Let $k$ be the number of connected components of $S$. \hlfix{Also $S$ is a spanning subgraph of the outer face of $t$ and $S$ has $n$ edges.}{Not sure what this means. There are 4 edges in the picture you sent.}  To join the components $S$ must have $k$ edges not in the outer face of $\tau(W)$.  Furthermore $t$ has $n-3$ edges not in its outer face and so also not in the outer face of $\tau(W)$.  Thus there are $n-3+k$ edges of $t$ not in the outer face of $\tau(W)$.

@@

\begin{cor}
For $W = (\cP, n)$ can be uniquely decomposed into maximal exact subdiagrams.
\end{cor}

\begin{proof}
This is a corollary to Lemma \ref{decompositionlem}, that gives a unique decomposition of polygon partitions, $\tau(W)$,  into maximal triangulated pieces, and Lemma \ref{lem triang to exact}, which gives a correspondence between triangulated components and exact subdiagrams.
\end{proof}

Furthermore, we can show that maximal triangulated pieces are disjoint in $\tau(W)$. \note{do we have this already?}

\begin{lem} \label{disjointexactsubdiagslem}
Let $W = (\cP, n)$ be an admissible Wilson loop diagram, and $(Q, V(Q))$, $(R, V(R))$ two maximal exact subdiagrams. Then $Q\cap R = \emptyset$.
\end{lem}

\begin{proof}
Consider an exact subset of a maximally exact subdiagram, $T \subset Q$. For $p \in Q \setminus T$, $|V(T \cup p)| = |T| +1 +3  = |V(T)| +1$. That is, by extending an exact subdiagram to a larger one, one extends the vertex set by exactly on vertex per proppagator. Therefore, if $T \neq \emptyset = Q \cap R$, \bas |V(Q \cup R)| = |T| +3 + |Q \setminus T| + |R \setminus T| = |Q \cup R| + 3\;.\eas Therefore, if $(Q, V(Q))$ and $(R, V(R))$ are two exact subdiagrams such that $Q\cap R \neq \emptyset$, then $(Q \cup R, V(Q \cup R))$ is exact.
\end{proof}

@@

\begin{proof}
One direction has been proved in previous work, but I give a different proof here to be consistent with the method of this document.

@@

Reorganize the vertex sets as follows: \ba \cup_{P_i \neq P_i'} V(P_i) = V(\cup_{P_i \neq P'_i} P_i)  \label{vertexsets}\; .\ea Since maximal exact subdiagrams are disjoint, and single propapagators may define maximal subdiagrams, $\cup_{i = 1}^k P_i = \cup_{i = 1}^k P'_i = \cP$. Then equation \eqref{vertexsets} becomes \bas \cap_{P_i \neq P_i'} (F(P_i^c))^c = F(\cup_{P_j = P'_j} P_j)^c \;.\eas These flats may, of course, be empty.

@@

It is a classical fact the the number of triangulations of an $n$-gon is the $n-2$ Catalan number, namely $\frac{1}{n-1}\binom{2(n-2)}{n-2}$.  Thus we can count the number of equivalent diagrams.

@@

\begin{figure}
  \includegraphics{specialp}
  \caption{The different possibilities for $D$ and $p$.  No other propagators can end in the fat red sections.  Other segments may have additional propagators ending in them.}\label{fig special p}
\end{figure}

@@

\begin{figure}
  \includegraphics{Le_diagrams}
  \caption{Le diagrams for $C$ (left) and $D$ (right).}\label{fig Le}
\end{figure}

@@

  \begin{figure}
    \includegraphics{messy}
    \caption{Plusses coming from $I_{n-2}^{C}$ (top left),  $I_{n-1}^{C}$ (top right) and $I_{n}^{C}$ bottom when $n-1, n-2\in I_{n-1}^{C}$.  The blue lines are the non-intersecting paths.  The dashed blue lines may or may not appear, but if one appears then they both do.}\label{fig messy}
  \end{figure}

@@

  \begin{figure}
    \includegraphics{messyD}
    \caption{Plusses coming from $I_{n-2}^{D}$.}\label{fig messyD}
  \end{figure}

>>>>

\begin{dfn}\label{WLdfn}
A Wilson loop diagram is given by the following data: a cyclicly ordered set $V$, along with a choice of first vertex, and $k$ pairs $\{p_r = (i_r, j_r)\}_{r=1}^k$ ordered such that $i_r +1 < j_r$. \end{dfn}

We depict this data as a convex polygon, with vertices labeled by $V$ (preserving the cyclic ordering), and $k$ wavy lines in the interior of the diagrams. We depict the choice of first vertex on $V$ by a $\bullet$. Note that the polygon gives the verices of $W$ a cyclic ordering, which the choice of first vertex gives it a compatible linear order. Both the cyclic and the linear order become the correct perspective at various points in this paper. The pair $p_r$ defines a wavy line from the $i_r^{th}$ edge (between $i_r$ and $i_r+1$) and the $j_r^{th}$ edge. These are called propagators.  The condition on $i_r$ and $j_r$ means that the propagator does not go between adjacent edges. Let $\cP = \{p_r\}_{r=1}^k$ be the set of propagators. Then we write \bas W = (\cP, V) \;.\eas

@@

\begin{dfn}\label{equivdfn} \todo{I feel like I need some sort of uniqueness of partition result to even be able to define this well. Or something. Can we discuss?}
There is an equivalence relationship on the set of admissible Wilson loops diagrams generated by the following binary relation: $W = (\cP, n) \sim W'= (\cP', n)$ if
\begin{enumerate}
\item There exist two different exact subdiagrams, $(P, V(P))$ and $(P', V(P'))$ of $W$ and $W'$ respectively such that $V(P) =  V(P')$.
\item The complementary subdiagrams are identical: $(\cP \setminus P, V(P)^c) = (\cP' \setminus P', V(P')^c)$.
\end{enumerate}
\end{dfn}

@@

The diagrams $W_1 \sim W_2$ because $(\{(5,8), (5,7), \{5,6,7,8,1\})$ and $(\{(5,8), (7,8), \{5,6,7,8,1\})$ are the corresponding differing subdiagrams. Furthermore, there is the equivalence $W_2 \sim W_3$ due to the exact subdiagrams $(\{(1,4), (3,4)\}, \{1,2,3,4,5\})$ and $(\{(1,4), (1,3)\}, \{1,2,3,4,5\})$. This forces an equivalence between $W_1$ and $W_3$, even though one cannot partition the propagators of each into a an exact subdiagram (that may vary between the diagrams) and a complement that is fixed.  \end{eg}

@@

\begin{dfn}
  Let $W = (\cP, n)$ be an admissible Wilson loop diagram.  The \emph{polygon partition} associated to $W$, denotes $\tau(W)$ is defined as follows.
  \begin{itemize}
  \item The vertices of $\tau(W)$ correspond to the edges of $W$.
  \item Labeling the vertices of $\tau(W)$ with the edge number of $W$gives a cyclic order to the vertices. Connecting consecutive vertices gives a graph theoretic cycle called the polygon of $\tau(W)$.
  \item Each propagator of $W$ defines a chord edge of $\tau(W)$ for each propagator of $W$.  If $p= (i,j) \in \cP$, it defines a chord connecting the verices $i$ and $j$
  \end{itemize}
\end{dfn}

\begin{lem}\label{tausimpleplanarlem}
If $W = (\cP, n)$ is an admissible Wilson loop diagram, $\tau(W)$ is a simple planar graph wihose outer face is a cycle. It is embedded such that the vertices all like on this infinite face. These vertices are cyclically ordered, with a coice of first vertex, giving it an additional compatible linear order.
\end{lem}

\begin{proof}
Since the vertices of $\tau(W)$ are labeled by the edges of $W$, which are cyclically ordered, this gives an ordering to the vertices. Furthermore, since the outer polygon of $W$ is a cycle, the outer face of $\tau(W)$ is also a cycle. Since $W$ is admissible, no pairs of propagators cross. Thereofore, it is a planar embedding. There are no propagators $p = (i, i+1)$. Therefore, there is exactly one edge connecting any two adjacent edges of $\tau(W)$. Similarly, there are not two propagators such that both $p$ and $q$ start and end at the edges $i$ and $j$. Therefore, no other two vertices of $\tau(W)$ can be connected by more than one edge. Finally, the embedding of $\tau(W)$ is induced from embedding of the graph $W$.
 \end{proof}

\begin{comment}
\hlfix{For an admissible $W$ no propagators cross and so no chord edges of $\tau(W)$ cross.  Thus, in graph theoretic language, we can view $\tau(W)$ as a planar embedding of a
graph with no cut vertices, where all vertices lie on the infinite face, along with a distinguished start vertex and direction around the infinite face.  From this viewpoint the polygon of $tau(W)$ is the boundary of the infinite face.}{I've smushed this together a bit in the prev. lem. Tell me if I've missed anything. Its a lemma for emphasis, not difficulty}
\end{comment}

@@

\begin{dfn}
  Let $W$ be an admissible Wilson loop diagram and $\tau(W)$ its polygon partition. A \emph{triangulated piece} of $\tau(W)$ is a 2-connected subgraph of $\tau(W)$ which is a triangulation. We will take the convention that a subgraph consisting of a single chord edge is called a \emph{trivial} triangulated piece.
\end{dfn}

@@

\begin{lem} \label{decompositionlem}
  For $W$ an admissible Wilson loop diagram, the polygon partition $\tau(W)$ has a unique decomposition into maximal triangulated pieces, and edges in the polygon of $\tau(W)$.
\end{lem}

@@

\begin{proof}
We begin by giving an algorithm for the decomposition, then prove its uniqueness. Let $W = (\cP, n)$, with $|\cP| = k$.

Let $T(W)$ be a the dual graph of $\tau(W)$ with the vertex corresponding the to the infinite face split. In other words, place a vertex on each finite face of $\tau(W)$. These vertices are connected if there is a chord edge separating the edges. Furthermore, if the face is bounded by an edge of the polygonal cycle of $\tau(W)$, the corresponding vertex gets a leaf edge for each such boundary. Since, by Lemma \ref{tausimpleplanarlem}, $\tau(W)$ is a simple planar graph, $T(W)$ is an uniquely defined graph.

We claim that $T(W)$ is a tree. Since $\tau(W)$ is a planar graph with $k$ chord edges and no internal vertices, there are $k+1$ internal faces of $\tau(W)$. Each of these internal faces correspond to a vertex of $T(W)$, and there are $n$ vertices of $T(W)$ from the splitting of the dual graph, we see that $T(W)$ has $n+k+1$ vertices. On the other hand, $T(W)$ has $n+k$ edges, corresponding to the $n +k$edts of $\tau(W)$. Therefore, $T(W)$is a tree.

By construction, each face of $\tau(W)$ is a polygon. Therefore, each vertex of $T(W)$ is at least 3 valent (one for each edge of the polygon). If split each vertex of $T(W)$ if it is strictly greater than $3$ valent. (In other words, we fail to split exactly when the face is a triangle). The connected components of $T(W)$ correspond to the decomposition of $\tau(W)$ into maximal triangulated pieces and edges orinignaly in the polygon of $\tau(W)$.  Let $f$ be the forest thus obtained. The vertices of $f$ either have valence 1, (if they correspond leaf vertices) or 3,  (if they correspond to triangulated faces). Subtrees with no trivalent vertices correspond to either edges in the polygon of $\tau(W)$, if they were originally leaves of $T(W)$, or to maximal trivial triangulations. Splitting at all the faces that are not triangles ensures maximality of the decomposition. If the splitting were not maximal, then one could add a triangle to a connected component of the splitting, but this would imply that that splitting happened at a valence $3$ vertex.

To see uniqueness, consider a different maximal decomposition of $\tau(W)$. This induces a splitting on $T(W)$, where each connected component of the new decomposition corresponds to a subtree. Call this forest $f'$. Since $f' \neq f$, there are two trees, $t$ and $t'$ in $f$ and $f'$ that are distinct, but share at least one edge of $T(W)$. Since $f'$ is also maximal, $t'$ is not a subtree of $t$. Therefore, the edges of $t'$ can be found in at least two trees in the forest $f$. In particular, there is a vertex $v$ in $t'$ that corresponds to a split vertex of $T(W)$ in the original decomposition. This implies that $v$ has valence greater than $3$ in $T(W)$, and thus the corresonding face of $\tau(W)$ is not a triangle. In otherwords, the decomposition corresponding to $f'$ is not a triangulation. \note{This is a variant of what you wrote in the e-mail. Let me know if you like it.}
\end{proof}

\begin{cor} \label{maxtriangdisjointcor}
Given a maximal decomposition of $\tau(W)$, the maximal triangulated pieces are edge disjoint.
\end{cor}

\begin{proof}
If two maximal triangulated pieces were not disjoint, then one could take the union of the edges involved to get a larger triangulated piece. \note{Is it this simple? This is a variant of the proof on propagators}
\end{proof}

@@

  Since $W$ is admissible, by lemma \ref{} $\tau(W)$ is a simple graph. Furthermore, $t$ is not all of $\tau(W)$.
  Let $t$ be a triangulated piece of the decomposition of $\tau(W)$ given in Lemma \ref{decomplem} and suppose $t$ has $n>2$ vertices.  We want to count how many edges of $t$ are not on the outer face of $\tau(W)$.  Consider the intersection of $t$ with the outer face of $\tau(W)$.  This is a possibly disconnected subgraph of the polygon of $\tau(W)$ with $n$ vertices, call it $S$.  Let $k$ be the number of connected components of $S$.   To join the components $S$ must have $k$ edges not in the outer face of $\tau(W)$.  Furthermore $t$ has $n-3$ edges not in its outer face and so also not in the outer face of $\tau(W)$.  Thus there are $n-3+k$ edges of $t$ not in the outer face of $\tau(W)$.

@@

\begin{cor} \label{uniqueproppartitioncor}
For $W = (\cP, n)$ can be uniquely decomposed into maximal exact subdiagrams. These subdiagrams maximal subdiagrams partition $\cP$.
\end{cor}

\begin{proof}
This is a corollary to Lemma \ref{decompositionlem}, that gives a unique decomposition of polygon partitions, $\tau(W)$,  into maximal triangulated pieces, and Lemma \ref{lem triang to exact}, which gives a correspondence between triangulated components and exact subdiagrams. Futhermore, corrollary \ref{maxtriangdisjointcor} shows that the maximal triangulations are edge disjoint, and thus the propagators defining maximal exact subsets are also disjoint. Combining this with the fact that by construction, the decomposition of $\tau(W)$ accounts for all chord edges, gives the desired unique partition of propagators.
\end{proof}

@@

\begin{proof}
One direction has been proved in previous work, but we give a different proof here to be consistent with the method of this document.

@@

Reorganize the vertex sets as follows: \ba \cup_{P_i \neq P_i'} V(P_i) = V(\cup_{P_i \neq P'_i} P_i)  \label{vertexsets}\; .\ea Since maximal exact subdiagrams partition $\cP$, by Lemma \ref{uniqueproppartitioncor}, write $cup_{i = 1}^k P_i = \cup_{i = 1}^k P'_i = \cP$. Then equation \eqref{vertexsets} becomes \bas \cap_{P_i \neq P_i'} (F(P_i^c))^c = F(\cup_{P_j = P'_j} P_j)^c \;.\eas These flats may, of course, be empty.

@@

Since there is a unique way to decompose $W$ into maximal exact subdiagrams, it is logical to ask how many diagrams there are in an equivalence class. It is a classical fact the the number of triangulations of an $n$-gon is the $n-2$ Catalan number, namely $\frac{1}{n-1}\binom{2(n-2)}{n-2}$.  Thus we can count the number of equivalent diagrams.

@@

\begin{figure}
 % \includegraphics{specialp}
  \caption{The different possibilities for $D$ and $p$.  No other propagators can end in the fat red sections.  Other segments may have additional propagators ending in them.}\label{fig special p}
\end{figure}

@@

\begin{figure}
%  \includegraphics{Le_diagrams}
  \caption{Le diagrams for $C$ (left) and $D$ (right).}\label{fig Le}
\end{figure}

@@

  \begin{figure}
 %   \includegraphics{messy}
    \caption{Plusses coming from $I_{n-2}^{C}$ (top left),  $I_{n-1}^{C}$ (top right) and $I_{n}^{C}$ bottom when $n-1, n-2\in I_{n-1}^{C}$.  The blue lines are the non-intersecting paths.  The dashed blue lines may or may not appear, but if one appears then they both do.}\label{fig messy}
  \end{figure}

@@

  \begin{figure}
    % \includegraphics{messyD}
    \caption{Plusses coming from $I_{n-2}^{D}$.}\label{fig messyD}
  \end{figure}
